---
title: "Explore edx data"
output: html_document
---

```{r}
# Load packages
require(dplyr)
require(ggplot2)
```

### Load the data
```{r}
data.path="/Users/kanhua/Documents/Big Data/edX open data 2013/"
rawfile="HMXPC13_DI_v2_5-14-14.csv"
#edxdata<-read.csv(paste(data.path,rawfile,sep=""))
#save(edxdata,file=paste(data.path,"HMXPC13_DI_v2_5-14-14.RData",sep=""))
load(paste(data.path,"HMXPC13_DI_v2_5-14-14.RData",sep=""))
edxdata$certified=as.logical(edxdata$certified)
edxdata$registered=as.logical(edxdata$registered)
edxdata$explored=as.logical(edxdata$explored)
```

### Load course details data:
```{r}
course.data.path="/Users/kanhua/Dropbox/Documents in Dropbox/Open Courseware/Udacity courseware/Data Analyst Nanodegree/Udacity DA projects/P3 Explore Data/data/course_details.csv"
course.data<-read.csv(course.data.path)
head(course.data)
```

### Join the data
```{r}
edxdata<-inner_join(x=edxdata,y=course.data,by="course_id")
```

Get dimension of data:
```{r}
dim(edxdata)
```
Show head and tail of the data
```{r}
head(edxdata)
tail(edxdata)
```

Get the columns
```{r}
names(edxdata)
```

Total number of unique users:
```{r}
dim(unique(edxdata$userid_DI))
```

The course names
```{r}
dim(unique(edxdata$course_id))
```


Get the number of user_DI
```{r}
users<-edxdata %>%
  group_by(userid_DI) %>%
  summarise(course_taken=n(),
                 total_certified=sum(certified))

users=inner_join(users,edxdata,by="userid_DI")
table(users$total_certified)
```

plot the histogram of registered users vs. LOE_DI
```{r}
ggplot(aes(LoE_DI),data=users)+geom_bar()
```

plot the histogram of registered users vs. gender
```{r}
ggplot(aes(gender),data=users)+geom_bar()
```
Other ideas: breakdown gender with course

plot the histogram of registered users vs. age
```{r}
users$age=2013-users$YoB
ggplot(aes(age,fill=course_taken),data=subset(users,users$age>0))+geom_bar(binwidth=5)
```



plot the histogram of certified users by ```LoE_DI```
```{r}
ggplot(aes(total_certified,fill=LoE_DI),data=subset(users,!is.na(users$LoE_DI)))+
  geom_bar()+
  scale_x_discrete()+
  scale_y_log10()
```

plot the histogram of certified users by ```gender```
```{r}
ggplot(aes(total_certified,fill=gender),data=subset(users,!is.na(users$gender)))+
  geom_bar()+
  scale_x_discrete()+
  scale_y_log10()
```

deal with age
```{r}
users$age=2012-users$YoB
table(users$age)
table(users$YoB)
```

#
```{r}
courses<-group_by(edxdata,course_id)
courses<-summarise(courses,joined_users=n(),
                   passed_num=sum(certified),
                   explored_num=sum(explored),
                   registered_num=sum(registered),
                   total_nforum_posts=sum(nforum_posts))
courses$pass.rate<-courses$passed_num/courses$registered_num
```

explored rate: passed/registered
```{r}
courses$pass.rate<-courses$passed_num/courses$registered_num
```

pass rate: passed/explored
```{r}
courses$hangon.rate<-courses$passed_num/courses$explored_num
```


```{r}
qplot(hangon.rate,data=courses)
unique(edxdata$LoE_DI)
unique(edxdata$gender)
```

### Participation of the courses
```{r}
course.stat<-group_by(edxdata,course_code)

course.stat<-summarise(course.stat,t_certified=sum(certified),
                       t_viewed=sum(viewed),
                       t_explored=sum(explored),
                       t_registered=sum(registered))

ggplot(aes(y=t_registered,x=course_code),data=course.stat)+geom_bar(stat="identity")
```

### Plot the distribution of nevents
```{r}

ggplot(aes(nevents),data=edxdata)+geom_freqpoly(aes(y=..density..,color=course_code))+xlim(0,1000)
```


### Plot the distrubution of nevents of certified users only
```{r}
ggplot(aes(nevents),data=filter(edxdata,certified==1))+geom_freqpoly(aes(y=..density..,color=course_code))+xlim(0,30000)
```

### Analyze average login events of certified users
```{r}
cert.edxdata <- filter(edxdata,certified == 1,!is.na(nevents))

cert.course.stat <- cert.edxdata %>%
  group_by(course_code) %>%
  summarise(mean_nevents = mean(as.double(nevents)),
            med_nevents = median(as.double(nevents))) %>%
  semi_join(course.data,by = "course_code") #use semi_join because we only want to get one result from course.data

cert.course.stat

```

### Plot the degree distribution of certified users
```{r}
ggplot(aes(LoE_DI),data=filter(edxdata,certified==1))+geom_histogram()+facet_wrap(~course_code)

#ggplot(aes(short_title,fill=LoE_DI),data=filter(edxdata,certified==1))+geom_histogram(position="fill")
```

### Plot the degree distribution of all users

```{r}

ggplot(aes(LoE_DI),data=edxdata)+facet_grid(course_code~.)+geom_histogram()

ggplot(aes(short_title,fill=LoE_DI),data=edxdata)+geom_histogram(position="fill")
```

### Extra: plot npost against nevents:

```{r}

ggplot(aes(x=ndays_act,y=nevents,color=certified),data=edxdata)+geom_point(position="jitter")+ylim(0,10000)

```

```{r}

ggplot(aes(nevents),data=edxdata)+geom_histogram()+xlim(0,10000)
```
