Exploring edX data during 2012 to 2013 by Kan-Hua Lee
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

require(dplyr)

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
```

```{r echo=FALSE, Load_the_Data,cache=TRUE}
# Load the Data
data.path = "./data/"
rawfile = "HMXPC13_DI_v2_5-14-14.csv"
edxdata<-read.csv(paste(data.path,rawfile,sep=""))

# Load course details data:
course.data <- read.csv(paste(data.path,"course_details.csv",sep=""))

# Join the edxdata and course.data
edxdata <- inner_join(x = edxdata,y = course.data,by = "course_id")

# Add a column "age"
edxdata$age = 2013 - edxdata$YoB

# Add education levels
levels(edxdata$LoE_DI) <-
c("","Less than Secondary","Secondary","Bachelor's","Master's","Doctorate")

# Add a new variable "access.days"
start.time <- strptime(edxdata$start_time_DI,"%Y-%m-%d")
last.event.time <- strptime(edxdata$last_event_DI,"%Y-%m-%d")
edxdata$access.period <- last.event.time - start.time

edxdata$access.period <-
as.numeric(edxdata$access.period,units = "days")

# Calculate access.rate
edxdata$access.rate <- with(edxdata,ndays_act / access.period)

edxdata$registered <- as.logical(edxdata$registered)
edxdata$explored <- as.logical(edxdata$explored)
edxdata$certified <- as.logical(edxdata$certified)
edxdata$viewed <- as.logical(edxdata$viewed)

```



```{r,echo=FALSE,group_by_user,cache=TRUE}
#Summarize the course statistics of each unique users
users <- edxdata %>%
  group_by(userid_DI,LoE_DI,age,gender) %>%
  summarise(
    course_taken = n(),
    total_registered = sum(registered),
    total_explored = sum(explored),
    user.certificates = sum(certified),
    user.age=mean(age)
  )

users$user.certificates = as.integer(users$user.certificates)
```


```{r, group_by_courses,echo=FALSE,cache=TRUE}
# Group the statistics of each unique course by course id
courses <- edxdata %>%
  group_by(course_id) %>%
  summarise(
    joined_users = n(),
    passed_num = sum(certified),
    explored_num = sum(explored),
    registered_num = sum(registered),
    total_nforum_posts = sum(nforum_posts),
    mean_nevents=mean(nevents)
  ) %>%
  inner_join(course.data,by = "course_id")

courses$pass.rate <- courses$passed_num / courses$registered_num

#The ratio between explored users and registered users
courses$hangon.rate <- courses$passed_num / courses$explored_num
```


```{r,group_by_users,echo=FALSE,cache=TRUE}
# Summarize the statistics of each unique course by each course. Essentially we merged the courses with the same course name
course.stat <- edxdata %>%
  group_by(course_code) %>%
  summarise(t_certified = sum(certified),
  t_viewed = sum(viewed),
  t_explored = sum(explored),
  t_registered = sum(registered)
)
```

# Univariate Analysis

### What is the structure of your dataset?

This dataset has 27 columns and 641,138 rows. Each row is the statistical results per user per course.
This dataset includes 16 courses. Three of those are the same course material that offered two different times. This dataset has 476,549 unique registrants of all courses. Detailed descriptions of this dataset can be found in [edx data summary.md](./edx data summary.md).

```{r,echo=FALSE,warning=TRUE}
dim(edxdata)
dim(users)
```

### What is/are the main feature(s) of interest in your dataset?

We are interested in the backgrounds and activities of the students and how these factors affect their performances in a course.
The features in this dataset that we will focus on are: ```registered```, ```explored```, ```LoE_DI```, ```YoB```, ```gender```, and ```nevents```.
The response that we are interested in is ```certified``` and ```grade```.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interested?
 
In this analysis we mainly use ```nevents``` as the main measure of how much efforts that a student makes for the course. Other features such as ```ndays_act```, ```start_time_DI```, ```last_event_DI``` and ```nforum_posts``` will occasionally be used to support this investigation.

### Did you create any new variables from existing variables in the dataset?

The following new variables were created in the original dataframe ```edxdata```:

  + ```age```: the age of the user when taking the course. It is calculated by 2013-```YOB```.
  + ```access.period```: Number fo days between ```last_event_DI``` and ```start_time_DI```.
  + ```access.rate```: ```ndays_act``` divided by ```access.period```. This variable measures how often an user accesses the course.

Also, we created the following new datasets by grouping certain features in the original dataset:

#### users

This dataset is created by grouping the raw dataset by ```userid_DI```. The following new variables were created:

  + ```total_registered```: number of courses registered.
  + ```total_explored```: number of courses explored.
  + ```user.certificates```: number of courses certified.

#### courses

This dataset is created by grouping the raw dataset by ```course_id```.
The following new variables were created:

  + ```passed_num```: total certified users of the course.
  + ```explored_num```: total users who explored the course.
  + ```registered_num```: total users who registered the course.
  + ```total_nforum_posts```: total number of posts in the course.
  + ```pass.rate``` : the number of certificated users divided by the number of registered users
  + ``` hangon.rate``` : the number of explored users divided by the number of registered users


### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

We did not found any unusual distributions to the best of our knowledge. 

In the other hand, we preprocessed a number of features of the raw dataset:

+ Transform the feature ```LoE_DI``` into levels.
+ Transform the features ```certified```, ```explored``` and ```viewed``` into logical data type.


# Univariate Plots Section

Firstly we investigate some basic user statistics of each course:

#### Number of registrants of each course:
```{r,echo=FALSE}
ggplot(aes(y = registered_num,x = abbrev_course_title),data = courses) +
  geom_bar(stat = "identity") + coord_flip()
```

#### Number of registrants who "explored" the course:
```{r,echo=FALSE}
ggplot(aes(y = explored_num,x = abbrev_course_title),data = courses) +
  geom_bar(stat = "identity") + coord_flip()
```

#### Number of registrants who earned certificates:

```{r,echo=FALSE,warning=FALSE}
ggplot(aes(y = passed_num,x = abbrev_course_title),data = courses) +
  geom_bar(stat = "identity")+coord_flip()
```

#### ```pass.rate``` of each course:
```{r,echo=FALSE,warning=FALSE}
ggplot(aes(x = abbrev_course_title,y = pass.rate),
       data = courses) +
       geom_bar(stat = "identity") +
       coord_flip()
```

From the above analysis, we can see that CS50X (Introduction to Computer Science I) is the most popular course in terms of the number of course registrants. Also, we can see that ```pass.rate``` of all courses are around a few percent.

We then investigate the statistics of the users:

#### Number of certificates earned by the registrants:
```{r,echo=FALSE,warning=FALSE}
ggplot(aes(user.certificates),data = users) +
  geom_bar() + scale_y_log10() +
  scale_x_discrete()
```
  
Please note that the y-axis of the above graph is in log scale.


#### Histogram of ```age``` among all registrants:
```{r,echo=FALSE,warning=FALSE,cache=TRUE}
ggplot(aes(age),
       data = subset(users,users$age > 0)) +
       geom_bar(binwidth = 5)
```
  
Although the distribution of age is very wide, most registrants are between 20 to 35 years old.

#### Histogram of ```LOE_DI``` of all registrants, with NA and blank ("") filtered:
```{r,LoE_DI_hist,echo=FALSE,warning=FALSE,cache=TRUE}
ggplot(aes(LoE_DI),
       data = filter(users,((LoE_DI != "") & (!is.nan(LoE_DI))))) +
       geom_bar()
```
  
```LoE_DI``` are dominated by *Less than Secondary*, *Master's* and *Doctorate*.


#### Histogram of ```gender``` of all registrants, with NA and blank ("") filtered:
```{r,gender_hist,echo=FALSE,warning=FALSE,cache=TRUE}
ggplot(aes(gender),
       data = filter(users,((gender != "") & (!is.nan(gender))))) +
  geom_bar()
```

We can see that these courses are dominated by male registrants.


Next, we investigates the activities of the registrants using the feature ```nevents```.

#### Histogram of ```nevents``` of all registrants:
```{r,echo=FALSE,warning=FALSE}
ggplot(aes(nevents),
       data = filter(edxdata, nevents <10000)) +
       geom_histogram() 
```

In this graph, we set the limit of x-axis to 10000 for clarity, since the data entries with ```nevents```>10000 are very rare. We can see that most of the registrants has ```nevents``` less than 1000.

#### Histogram of ```nevents``` of registrants that passed a course (```certified==1```)
```{r,echo=FALSE,warning=FALSE}
ggplot(aes(nevents),
       data = filter(edxdata, certified == 1 & nevents >0)) +
       geom_histogram() 
```

Within the group of ```certified==1```, the distribution of ```nevents``` becomes more evenly distributed between 0 and 10000.

# Bivariate Plots Section

### Investigating the background of the registrants

#### Gender distribution of all registrants, by course
```{r,echo=FALSE,warning=FALSE}
ggplot(aes(abbrev_course_title,fill = gender),data = edxdata) +
  geom_bar() + coord_flip()
```
  
The figure shows that most of the courses are dominated by male registrants.


#### Gender distribution of certificated registrants, by course
```{r,echo=FALSE,warning=FALSE}
ggplot(aes(abbrev_course_title,fill = gender),data = filter(edxdata,certified == 1)) +
  geom_bar() + coord_flip()
```

The certificated registrants of most courses are also dominated by men, except *Poverty* and *HealthStat*.

#### The distribution ```LoE_DI```, by course:
```{r,LoE_DI_distri,echo = FALSE,warning = FALSE,cache = TRUE,fig.width = 9.5,fig.height =7}
ggplot(aes(LoE_DI,fill = certified),data = edxdata) +
  geom_histogram() +
  facet_wrap( ~ abbrev_course_title) +
  scale_y_log10() +
  coord_flip()
```
  
The above figure shows that the population of registrants with *Bachelor's* and *Secondary* degree are relatively small in all courses. Note that the x-axis is in log scale.


### Investigating the activities of registrants

#### Normalized distribution of ```nevents``` of all registrants, by course
```{r,echo=FALSE,warning=FALSE}
ggplot(aes(nevents),
       data = filter(edxdata,!is.na(nevents))) +
       geom_freqpoly(aes(y = ..density..,color = abbrev_course_title)) +
       xlim(0,1000) +
       scale_y_log10()
```
  
The distribution of most of the courses are very similar. All the curves drop very sharply below 100 ```nevents```. After that, the decrease becomes more moderate. 

#### Distrubution of ```nevents``` of all registrants with certificates, by course
```{r,echo=FALSE,warning=FALSE}
ggplot(aes(nevents),
       data = filter(edxdata,certified == 1)) +
  geom_freqpoly(aes(y = ..density..,color = abbrev_course_title)) +
  xlim(0,30000)
```
  
The distribution of ```nevents``` becomes more Gaussian if we only consider the population of those with certificates. However, the peaks of these distributions vary from course to course. 

#### Boxplot of ```nevents``` of registrants with certificates, by course
```{r,echo=FALSE,warning=FALSE}
ggplot(aes(x = abbrev_course_title,y = nevents),
       data = filter(edxdata,certified == 1)) +
       geom_jitter(alpha = 0.2) +
       ylim(0,30000) +
       geom_boxplot() +
       coord_flip()
```

This plot is a different representation of the previous plot. This boxplot shows the median and variations of ```nevents``` more clearly. Although the median of values of ```nevents``` of each course vary a lot, most courses have median ```nevents``` values around a few thousands. The course *EM* has the highest median value of ```nevents```, while the course *CSH* has the lowest value.

In the next three graphs, we plot similar boxplots of ```ndays_act```, ```access.period``` and ```access.rate``` instead of ```nevents```. 

#### Boxplot of ```ndays_act``` of registrants with certificates, by course
```{r,echo=FALSE,warning=FALSE}
ggplot(aes(x = abbrev_course_title,y = ndays_act),
       data = filter(edxdata,certified == 1)) +
       geom_jitter(alpha = 0.2) +
       geom_boxplot() +
       coord_flip()
```

#### Boxplot of ```access.period``` of registrants with certificates, by course
```{r,echo=FALSE,warning=FALSE}
ggplot(aes(x = abbrev_course_title,y = access.period),
       data = filter(edxdata,certified == 1)) +
       geom_jitter(alpha = 0.2) +
       geom_boxplot() +
       coord_flip()
```

#### Boxplot of ```access.rate``` of registrants with certificates, by course
```{r,echo=FALSE,warning=FALSE}
ggplot(
  aes(x = abbrev_course_title,y = access.rate),
  data = filter(edxdata,certified == 1 &
  access.rate > 0 & access.rate < 1)
  ) +
  geom_jitter(alpha = 0.2) +
  geom_boxplot() +
  coord_flip() +
  ylim(0,1)
```

We can see that ```ndays_act```, ```access.period``` and ```access.rate``` all vary a lot between each course, as we observed in the boxplot of ```nevents```.

We then pick two user activities features ```nevents``` and ```ndays_act```, and explore their relation:

#### ```ndays_act``` against ```nevents``` among the "explored" registrants (```explored==1```):
```{r,echo=FALSE,warning=FALSE}
ggplot(
  aes(x = ndays_act,y = nevents,color = certified),
  data = filter(edxdata,explored == 1 &
  !is.na(nevents) & !is.na(ndays_act))
  ) +
  geom_jitter(alpha = 0.5,size = 1) +
  ylim(0,20000)
```

From this plot, we can see that ```n_days_act``` and ```nevents``` have strong correlation. Also, data entries with higher values of ```nevents``` and ```n_days_act``` have higher chance to be certified.

In the last part of this section, we explore some other features.

#### ```total_nforum_posts``` against ```pass.rate``` of each course
```{r,echo=FALSE,warning=FALSE}
ggplot(aes(x = total_nforum_posts,y = pass.rate,
           color = abbrev_course_title),data = courses) +
           geom_point()
```

Since ```total_nforum_posts``` might be an indicator of support offered by the community. One may assume that large amount of ```total_nforum_posts``` may therefore help the ```pass.rate```. However, the we did not see such relation in the above plot.


# Bivariate Analysis


### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

1. In every course except *Proverty*(The Challenges of Global Proverty), registrants with certificates are dominated by male users.

2. Registrants with certificates are largely dominated by those who do not have a Secondary school degree (*less than Secondary*) or those who hold an advanced degree (*Master* and *PhD*). This trend exists in every course in this dataset.

3. The histogram of ```nevents``` is exponentailly distributed among all registrants, but the distribution becomes more Gaussian when only certificated registrants are included.

4. The variation of ```nevents``` among certificated users of each course is large. The average of ```nevents``` of each course is typically around a few thousand times.

5. The course *EM* (Electricity and Magnetism) has highest mean values of ```nevent```, ```access.rate``` and ```ndays_act``` across all courses, which suggests that it may be the most demanding course within these 16 courses.

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

1. ```total_nforum_posts``` is not strongly correlated to ```pass.rate```, indicating that the support from the community in the forum is not an essential factor for ```pass.rate```.

2. Most users has 100 ```nevents``` per ```n_day_act```, as shown in the figure of ```nevents``` against ```n_day_act```. It also shows that more ```nevents``` and ```n_day_act``` has higher chance to pass the course.

### What was the strongest relationship you found?

Within these 16 courses, registrants who passed these courses are dominated by male users. Also, the certificated students are dominated by those do not have a Secondary school degree (*less tha Secondary*) or hold an advanced degree (*Master* and *PhD*).


# Multivariate Plots Section

In this section, we will explore the activities of users versus ```grade```.

#### Distribution of ```grade``` of each course
```{r,echo=FALSE,warning=FALSE,cache=TRUE,fig.width=9}
ggplot(
  aes(grade,fill=certified),
  data = filter(edxdata,explored==1)
  ) + geom_histogram(aes(y=..density..))+ ylim(0,15)+
  facet_wrap(~abbrev_course_title)
```

The distribution of ```grade``` is different from course to course, but almost all courses have a peak near zero, as expected. Note that ```grade``` of the course ```CS50X``` is not available.

#### ```nevents``` against ```grade```, by course
```{r,echo=FALSE,warning=FALSE,cache=TRUE,fig.width=9.5}
ggplot(aes(x = nevents,y = grade,color=certified),
       data = filter(edxdata,explored == 1)) +
  geom_point(size=0.5) + facet_wrap(~abbrev_course_title) +xlim(0,20000)
```


#### ```access.rate``` against ```grade```, by course
```{r,echo=FALSE,warning=FALSE,cache=TRUE,fig.width=10}
ggplot(aes(x = access.rate,y = grade,color=certified),
       data = filter(edxdata,explored == 1)) +
  geom_point(size=0.5) + xlim(0,1)+ facet_wrap(~abbrev_course_title)
```

The above figures show that the correlation between ```grade``` and ```nevents``` or ```access.rate``` are not very strong.


# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

From these figures, we found that the correlation between ```grade``` and ```nevents``` or between ```grade``` and ``` access.rate``` are not very strong. However, courses such as *EM*  or *CSM13* does show some weak trends that more ```access.rate``` or ```nevents``` results in higher ```grade```.

### Were there any interesting or surprising interactions between features?

The distribution of ```grade``` behaves differently from course to course. Two different patterns are commonly seen:

1. M-shape: Two main peaks exist in the distribution. One is in the region below the pass grade, while another one is in the region above pass grade. *Biology*, *CSM12*, *CSM13*, *Poverty* has M-shape distribution.

2. U-shape: Most of the population accumulated at both ends of the range. The distribution of *Circuits12*, *Circuits13* and *JusticeX* has this characteristic.


### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, plot_one,cache=TRUE}
ggplot(aes(abbrev_course_title,fill = LoE_DI),
       data = filter(edxdata,((LoE_DI!="") & (!is.nan(LoE_DI))))) +
  geom_histogram(position = "fill") + 
  coord_flip()+
  facet_wrap(~certified)+
  scale_fill_discrete(name="Level of education") + 
  xlab("course title (abbreviated)") +
  ylab("normalized population of registrants") +
  ggtitle("Density plot of registrants' level of education by course and certification status")
  
```

### Description One
This figure is the density plot of education level (```LOE_DI```) of the registrants by each course, with and without certificates. *TRUE* are the registrants with certificates, whereas *FALSE* are the registrants without certificates. Data entries with ```LOE_DI==NA or ""``` are ignored. As shown in this figure, the registrants of each course are dominated by the participants who hold an advanced degree or less than Secondary degree. The population of users who hold a secondary or Bachelor's degree are small in every course. In addition, this plot shows that the composition of education levels does not vary a lot bewteen the population with and without certificates.

### Plot Two
```{r,echo = FALSE, plot_two, warning = FALSE,cache = TRUE}
ggplot(aes(x = abbrev_course_title,y = nevents),
       data = filter(edxdata,certified == 1)) +
       geom_jitter(aes(color = short_title),alpha = 0.2,na.rm = TRUE) +
       ylim(0,30000) +
       geom_boxplot() +
       coord_flip() +
       xlab("course title (abbreviated)") +
       ylab("nevents:number of interactions with the course (times)") +
       theme(legend.position = "none") +
      ggtitle("Boxplot and scatter plot of certified registrants by course")
```

### Description Two

This figure shows a box plot and a scatter plot of ```nevents``` of each certified participant versus each course. The upper, middle and lower hinges of the boxplot represent the 25th, 50th 75th quantiles of```nevents```,respectively. ```nevents``` records the number of interactions with the course and therefore can be an indication of the efforts required to pass a course. The boxplot helps comparing the median and spread of ```nevents``` between each course. For example, the median and variation of ```nevents``` of the course *CSH* is very small compared to other courses. On the other hand, the course *EM* has the highest median ```nevents``` value among courses, indicating this course needs more efforts to complete.


### Plot Three
```{r,echo=FALSE,warning=FALSE, plot_three, cache=TRUE}
ggplot(aes(x = nevents/10000,y = grade,color=certified),
       data = filter(edxdata,explored == 1)) +
  geom_point(size=0.5) + 
  facet_wrap(~abbrev_course_title) +
  xlim(0,2) +
  xlab("nevents:number of interactions with the course (10000 times)") +
  ylab("grade (rescaled to 0~1)") +
  ggtitle("Grade against nevent among the registrants who \"explored\" the course")

```

### Description Three

The figure shows ```grade``` against ```nevent``` of each course among "explored" registrants (```explored==1```). Different colors in the plot separate the registrants who earned the certificates or not. This set of figures shows the correlation between ```grade``` and ```nevent``` are weak. However, courses such as *CSM13* or *EM* shows slightly stronger correlation between ```grade``` and ```nevent```. 

------

# Reflection
In this study we analyzed which and how the background information and activities of users affect their performance in a MOOC course. In most of the analysis, we mainly choose the "explored" registrants (```explored==1```) as the sample space because including all the registrants of each course may make the results very biased because most of the registrants do not involve a lot in the courses. However, it is worth noting that a few percent of certificated registrants are not "explored" registrants. These extreme cases are neglected in our analyses. In addition, we choose to use ```certified``` and ```grade``` to gauge the performance of the registrants, but different grading policies and requirements for certificates make it difficult to find universal trends between each course.

Through these data visualizations, we found some significant trends in gender and level of education among the registrants who earned certificates. However, we struggled to find strong correlations between the performance against features such as ```nevents```, ```n_days_act``` or ```access.rate```, because all the participants have different level of background and targets for exploring a course, which is also an attractive nature of MOOC.

It is possible to explore this dataset further by adopting some statistical learning methods such as regressions or decision trees to predict ```certified``` or ```grade```. Also, the correlations between users' countries (```final_cc_cname_DI```) and other features would also be interesting to look into.
