---
title: "Data Analysis with R—Problem Set 5"
output: html_document
---
Load the packages
```{r}
library(ggplot2)
data(diamonds)
?diamonds
```

Get the dimension and summary
```{r}
dim(diamonds)
summary(diamonds)
```

```{r}
summary(diamonds$color)
```


```{r}
qplot(price, data = diamonds)
mean(diamonds$price)
median(diamonds$price)
```

# Q1
```{r,echo=TRUE}
ggplot(aes(price,fill=cut),data=diamonds)+geom_bar()+facet_wrap(~color)+scale_fill_brewer(type = 'qual')
```

### Q2: Create a scatterplot of diamond price vs. table and color the points by the cut of the diamond.
```{r}
ggplot(aes(x=table,y=price,color=cut),data=diamonds)+
  geom_point(aes(color=cut))+scale_color_brewer(type='qual')
```
Note: in Q1, we use ```scale_fill_brewer``` to set the theme of color, but in Q2 we use ```scale_color_brewer```.


### Q4:
Create a scatterplot of diamond price vs. volume (x * y * z) and color the points by the clarity of diamonds. Use scale on the y-axis to take the log10 of price. You should also omit the top 1% of diamond volumes from the plot.

Note: Volume is a very rough approximation of a diamond's actual volume.

Note: In the link, a color palette of type 'div' was used to color the scatterplot using ```scale_color_brewer(type = 'div')```

Note: ```quantile(diamonds$vol,0.99)``` is different from ```0.99*max(diamonds$vol)```.
The former takes the last 99% of the rank, while the later one takes 0.99 of the maximum value.

```{r}
diamonds$vol = with(diamonds,x * y * z)
s_diamonds = subset(diamonds,vol < quantile(diamonds$vol,0.99))
ggplot(aes(x = vol,y = price,color = clarity),data = s_diamonds) +
  geom_point() +
  scale_y_log10() +
  scale_color_brewer(type = 'div')
```

### Extra: Density plot
examples from ggplot book p.69.
We can use ```y=..density``` to normalize the bin counts and presnets the density.
```{r}
depth_dist <- ggplot(diamonds,aes(depth)) + xlim(58,68)
depth_dist + geom_histogram(aes(y = ..density..),binwidth = 0.1) + 
  facet_grid(cut ~.)
depth_dist + geom_histogram(aes(fill = cut),binwidth = 0.1,
                            position = "fill")
depth_dist + geom_freqpoly(aes(y = ..density..,color = cut),
                           binwidth = 0.1)

```

